---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(data.table)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# bidsr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/dipterix/bidsr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dipterix/bidsr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

Brain Imaging Data Structure (`BIDS`) is a standard for organizing neuroimaging and behavioral data (see https://bids.neuroimaging.io/index.html). 
Package `bidsr` aims at providing comprehensive tools to query and manipulate 'BIDS' data files. 
This package is experimental, please see [Road-map](#road-map) below.

## Installation

You can install the development version of `bidsr` like so:

``` r
# install.packages("pak")
pak::pak("dipterix/bidsr")
```

## Road-map

Current milestone:

* Base classes
  - [X] R `S7` base classes and properties
  - [X] tabular data
  - [X] `JSON` side-car
  - [X] `BIDS` entity
* Class definitions for common files
  - Modality agnostic files:
    * [X] `dataset_description.json`
    * [X] `README`
    * [ ] `CITATION.cff`, `CHANGES`, `LICENSE`
    * [X] `participants.tsv`, `participants.json`
    * [X] `samples.tsv`, `samples.json`
    * [X] `phenotype/`
    * [X] Scans file
    * [X] Sessions file
  - Modality specific files:
    * [X] Query files by data type for each subject
    * [X] Use `BIDS` schema to generate file entity rules dynamically
* Query `BIDS` files
  - [X] Get top-level files
  - [X] Analyze and list subject data by types
  - [ ] Find side-car and meta-data information using the `BIDS` inheritance principle (test needed)

Next milestone:

* Provide read functions to handle common file formats
  - [X] `JSON` format
  - [X] `.tsv` or `.csv` format
  - [ ] `.nii` format
  - [ ] `.mat`, `.edf`, ... (electrophysiology)
* [ ] Provide handler registry to access and process data files

A tentative to-do list:    

* [ ] Modify the path
* [ ] Use `BIDS` schema to validation rules dynamically
* [ ] Validate `BIDS` files

## Example

This example demonstrates high-level tools to query a 'BIDS' project. 
The code requires executing a built-in command to download ['BIDS' example data-sets](https://github.com/bids-standard/bids-examples). 
Please copy-paste-run the following R commands:

```{r, results='hide'}
library(bidsr)
example_root <- download_bids_examples()
```

Let's inspect the project `ieeg_motorMiller2007` from the official `BIDS` example repository:

```{r}
## Path to `ieeg_motorMiller2007`
project_path <- file.path(example_root, "ieeg_motorMiller2007")

## BIDS project instance
project <- bids_project(path = project_path)

## The top-level `dataset_description.json` and `participants.tsv` 
## can be parsed via
description <- project$get_bids_dataset_description()
participants <- project$get_bids_participants()


## To obtain the values, use `$` operator (same as `.` in `Python`):
description$BIDSVersion

## Given a project path or instance, a `BIDS` subject can be instantiated via 
## the following two ways. 

# Method 1
subject <- bids_subject(project = project, subject_code = "sub-bp")

# Method 2
subject <- bids_subject(project = project_path, subject_code = "bp")

## The official `BIDS` and its extensions support various of data types, 
## such as `anat`, `func`, `meg`, `eeg`, `ieeg`, etc. 
## use `query_modality` method to query the files
ieeg_table <- query_bids(subject, "ieeg")
print(ieeg_table)


## The first column contains a list of file instances. Here's an 
## example to read the `ACPC` electrode coordinates:
subset_result <- subset(ieeg_table, space == "ACPC" & suffix == "electrodes")
print(subset_result)

# Get parsed 'BIDS' file object
path_parsed <- subset_result$parsed[[1]]
print(path_parsed)

electrode_path <- file.path(project, path_parsed)
as_bids_tabular(electrode_path)


## `BIDS` entity

## Object `path_parsed` is a file entity class, with the entities parsed
class(path_parsed)

path_parsed$get_bids_entity("space")

## List all available entities
path_parsed$entities

## If supported by schema, the `BIDS` entity rules for the 
## file can be queried via
path_parsed$get_bids_entity_rules()
```




